### 2019-03-28 IEEE_TC_submission.d/symmetry_properties.tex
ssr

\begin{observation}\label{obs:rsym_canalyzing}
Suppose $f(x_1, x_2, \ldots, x_n)$ is an $r$-symmetric Boolean function.
Consider any variable $x_i$ and suppose
the number of variables in the group
containing $x_i$ is $\nu_i$, $1 \leq i \leq n$.
\begin{enumerate}
\item Function $f$ is canalyzing in variable $x_i$
with canalyzing value 1 iff all the table entries where the number
of 1's in the group containing $x_i$ is nonzero have the same value
of $f$. (This value of $f$ is the canalyzed value when $x_i = 1$.)
\item Function $f$ is canalyzing in variable $x_i$ with
canalyzing value 0 iff all the table entries where the number of
1's in the group containing $x_i$ is less than $\nu_i$ have the same
value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
\QED
\end{enumerate}
\end{observation}
We now explain how Observation~\ref{obs:rsym_canalyzing} can be
used to develop an iterative algorithm for determining whether 
$f$ is an NCF; if so, the algorithm also constructs a default-normalized 
representation for $f$.
We use the following notation.
At beginning of iteration $j$, $X_j$ denotes the set of remaining variables,
$T_j$ denotes the table which provides values for the function,
$\mu_j$ denotes the number of rows of $T_j$ and $\ell_j$ 
denotes the number of remaining groups.
Initially (i.e., $j = 1$), $X_1$ = $X$ = $\{x_1, \ldots, x_n\}$,
$T_1 = T$ (the given table $T$ for $f$), 
$\mu_1 = \mu$ (the number of rows of $T$) and $\ell_j = r$ (the number of
symmetry groups of $f$); further, the NCF representation for $f$ is empty.
The algorithm carries out iteration $j$ if $\ell_j \geq 1$; 
in that iteration, the algorithm performs Steps I, II and III
as described below.

@@

\smallskip
\noindent
II. If yes, perform the following steps. 
\begin{enumerate}
\item Let $\alpha$ and $\beta$ be the respective canalyzing and 
canalyzed values for $x_i$ found in Step~I. 
\item For each variable $x_p$ in the group containing $x_i$, append the rule~
``$x_p\,:\,\alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
\item Obtain $X_{j+1}$ by deleting from $X_j$ all the variables which are 
in the same group as $x_i$.
\item Obtain $T_{j+1}$ by retaining only those rows of $T_j$ where all the
variables in the group containing $x_i$ are set to $\overline{\alpha}$. 
\item Set $\ell_{j+1} = \ell_j - 1$.
\item If $\ell_{j+1} \geq 1$, start the next iteration; 
      otherwise, \textbf{stop}.
\end{enumerate}

@@

To estimate the running time, we note that in iteration $j$, we can
determine whether there is a canalyzing variable in $X_j$
in $O(\ell_j \mu_j)$ = $O(r \mu_j)$ time using 
Observation~\ref{obs:rsym_canalyzing}.
(This is done by considering an arbitrary variable 
from each group.)
Once a canalyzing variable is identified, the other steps in that iteration
can be completed in $O(\mu_j)$ time.
Thus, the time for iteration $j$ is $O(r \mu_j)$.
We note that in each iteration, the number of rows in the table
is reduced by a factor of at least 2; that is, $\mu_{j+1} \leq \mu_j/2$.
To see why, suppose the canalyzing variable $x_i$ found in iteration $j$
is in group $g_p$ with $m_p$ variables. 
Thus, in $T_j$, there are $m_p+1 \geq 2$ possibilities for the number of 1's 
in group $g_p$. 
As indicated in Substep~4 of Step~II above, $T_{j+1}$ 
contains only those rows of $T_j$ where all the variables 
in $g_p$ have the same value.
In other words, there is only one possibility for the number of
1's in group $g_p$.
Thus, $\mu_{j+1}$, the number of rows in $T_{j+1}$,
is at most $\mu_j/(m_p+1)$ $\leq$ $\mu_j/2$.
It follows by simple induction that $\mu_j \leq \mu/2^{j-1}$.
Thus, the running time over all the $r$ iterations is given by
$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$, since the
geometric sum is bounded by $2\mu$.
The following theorem summarizes this result.

>>>>

\begin{observation}\label{obs:rsym_canalyzing}
Suppose $f(x_1, x_2, \ldots, x_n)$ is an $r$-symmetric Boolean function.
Consider any variable $x_i$ and suppose
the number of variables in the group
containing $x_i$ is $\nu_i$, $1 \leq i \leq n$.
\begin{enumerate}
\item Function $f$ is canalyzing in variable $x_i$
with canalyzing value 1 iff all the table entries where the number
of 1's in the group containing $x_i$ is nonzero have the same value
of $f$. (This value of $f$ is the canalyzed value when $x_i = 1$.)
\item Function $f$ is canalyzing in variable $x_i$ with
canalyzing value 0 iff all the table entries where the number of
1's in the group containing $x_i$ is less than $\nu_i$ have the same
value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
\QED
\end{enumerate}
\end{observation}
We now explain how Observation~\ref{obs:rsym_canalyzing} can be
used to develop an iterative algorithm for determining whether 
$f$ is an NCF; if so, the algorithm also constructs a default-normalized 
representation for $f$.
We use the following notation.
At beginning of iteration $j$,~
$\ell_j$ denotes the number of remaining groups,
$X_j$ denotes a set of variables with
exactly one variable from each remaining group,
$T_j$ denotes the table which provides values for the function and
$\mu_j$ denotes the number of rows of $T_j$.
Initially (i.e., $j = 1$),~ $\ell_1 = r$ (the number of symmetry groups of $f$),
$X_1$ is constructed by choosing one 
variable (arbitrarily) from each of the $r$ groups,
$T_1 = T$ (the given table $T$ for $f$) and 
$\mu_1 = \mu$ (the number of rows of $T$);
further, the NCF representation for $f$ is empty.
The algorithm carries out iteration $j$ if $\ell_j \geq 1$; 
in that iteration, the algorithm performs Steps I, II and III
as described below.

@@

\smallskip
\noindent
II. If yes, perform the following steps. 
\begin{enumerate}
\item Let $\alpha$ and $\beta$ be the respective canalyzing and 
canalyzed values for $x_i$ found in Step~I. 
\item For each variable $x_p$ in the group containing $x_i$, append the rule~
``$x_p : \alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
\item Set $X_{j+1}$ = $X_j - \{x_i\}$.  
\item Obtain $T_{j+1}$ by retaining only those rows of $T_j$ where all the
variables in the group containing $x_i$ are set to $\overline{\alpha}$. 
\item Set $\ell_{j+1} = \ell_j - 1$.
\item If $\ell_{j+1} \geq 1$, start the next iteration; 
      otherwise, \textbf{stop}.
\end{enumerate}

@@

To estimate the running time, we note that in iteration $j$, we can
determine whether there is a canalyzing variable in $X_j$
in $O(\ell_j \mu_j)$ = $O(r \mu_j)$ time using 
Observation~\ref{obs:rsym_canalyzing}.
(This is done by considering each variable in $X_j$.)
Once a canalyzing variable is identified, the other steps in that iteration
can be completed in $O(\mu_j)$ time.
Thus, the time for iteration $j$ is $O(r \mu_j)$.
We note that in each iteration, the number of rows in the table
is reduced by a factor of at least 2; that is, $\mu_{j+1} \leq \mu_j/2$.
To see why, suppose the canalyzing variable $x_i$ found in iteration $j$
is in group $g_p$ with $m_p$ variables. 
Thus, in $T_j$, there are $m_p+1 \geq 2$ possibilities for the number of 1's 
in group $g_p$. 
As indicated in Substep~4 of Step~II above, $T_{j+1}$ 
contains only those rows of $T_j$ where all the variables 
in $g_p$ have the same value.
In other words, there is only one possibility for the number of
1's in group $g_p$.
Thus, $\mu_{j+1}$, the number of rows in $T_{j+1}$,
is at most $\mu_j/(m_p+1)$ $\leq$ $\mu_j/2$.
It follows by simple induction that $\mu_j \leq \mu/2^{j-1}$.
Thus, the running time over all the $r$ iterations is given by
$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$, since the
geometric sum is bounded by $2\mu$.
The following theorem summarizes this result.
